\documentclass{article}

\usepackage[english]{babel} % Multilanguage support for headings and such
\usepackage[backend=biber]{biblatex} % Bibliography support using biber
\usepackage[margin=32mm]{geometry} % Allows for customisation of margins and sizes
\usepackage[hidelinks]{hyperref} % Creates clickable links (+ hide the red box)
\usepackage{csquotes} % Improves the looks of quotes `' and ``'' (used by biblatex references)
\usepackage{tikz} % Allows fancy diagrams
\usepackage{graphicx} % Adds many graphic support tweaks

% Adds and improves mathematical notation
\usepackage{physics}
\usepackage{amsmath}
\usepackage{amssymb}


\addbibresource{main.bib}

\newcommand{\Exp}[1]{\big\langle #1 \big\rangle}
% some TikZ style shortcut definitions
\tikzstyle{order4}=[
    circle,
    inner sep = 2pt,
    draw = green,
    fill = green,
]
\tikzstyle{maybe_order4} = [
    circle,
    inner sep = 2pt,
    draw = green,
    fill = white,
]
\tikzstyle{triangle}=[
    draw = #1,
    fill = #1!50,
    ultra thick,
]
\tikzstyle{edge} = [
    draw = #1,
    ultra thick
]
\interfootnotelinepenalty=10000

\title{{\huge 2D Causal Dynamical Triangulations} \\ {\Large using Markov-chain Monte Carlo methods}}
\author{\textit{Institute for Mathematics, Astrophysics and Particle Physics, Radboud University} \\[1mm] T.B.H. Gerstel \hspace{1cm} J.G.R. van der Duin}

\begin{document}

\maketitle

\begin{abstract} % (including research question and main results)
    Causal Dynamical Triangulation is a method to regularise the gravitational path integral by building a piece-wise flat spacetime from simplices.
    The goal of this project is to create an efficient 2D numerical model which can generate these piece-wise flat spacetimes or triangulations, and to use these to measure observables in 2D CDT using Monte Carlo methods to verify the validity of the model by comparison with theory.
    The simulational model is created in Rust and uses a Markov-chain to generate random triangulations. Data analysis is done in Python.
    Using this model, we are able to measure the length covariance on a two-dimensional torus for different triangulation sizes of up to 100,000 triangles and demonstrate that the theoretical expectation is within the statistical error our results.
\end{abstract}

\newpage

\section{Introduction}
% Introduction to the research area and research question
\input{1introduction.tex}

\section{Methods}
% explaining the algorithms and why they work
\input{2methods.tex}

\section{Results and Discussion}
% Including plots and final quantitative outcomes with errors
\input{3results_discussions.tex}

\section{Conclusion}
% Conclusions
\input{4conclusions.tex}

\printbibliography

\appendix
\section{Individual Contributions}
% explain briefly how you divided the work among each other on the various aspects: coding, data collection, data analysis, report writing
For the project we divided the work as follows:
\begin{itemize}
    \item \textit{Literature research}: Is done roughly equally
    \item \textit{Model design and programming}: Tom started with a vertex-based model, Jesse started a triangle-based model, and we optimized that together. Tom also started working on a shard model, but this was quickly discarded. Tom designed the CLI, and Jesse the visualisation.
    \item \textit{Data analysis}: Jesse performed most of the data analysis, and collected all the data for the preliminary analysis. Tom collected the data for the main measurements and verified the validity of the data analysis.
    \item \textit{Report writing}: Writing is divided roughly equally, generally Tom was responsible for the Introduction and Methods, and Jesse for the Results with Discussion.
\end{itemize}
For a very detailed overview of the individual contributions, they can be found in the commits of the GitHub repository, see appendix \ref{sec:coderepo}

\section{Code and Data distribution} \label{sec:coderepo}
For this project we used a public GitHub repository (see \url{https://github.com/SirBlueRabbit/monte-carlo-CDT}) to share all project files and all relevant data.
For an overview of the contents of the repository see the \textsc{README}.


\end{document}